Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2024-05-08 manuscript/manuscript.tex
edits/comments on simulation

\documentclass[12pt, titlepage, letterpaper]{article}
%\documentclass[12pt, letterpaper, titlepage]{article}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{setspace}

@@

\title{Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test}
  \author{Mathew Chandy %\\
%   \href{mailto:mathew.chandy@uconn.edu}
%   {\nolinkurl{mathew.chandy@uconn.edu}}\\
%  Elizabeth Schifano\\
%  Jun Yan\\
%  Xianyang Zhang\\
}

@@

Another prevalent misapplication of the KS test arises when data exhibit serial
dependence, which is often overlooked in analyses. For example, in an
investigation of performance variation in high-performance computing systems,
\citet{tuncer2019ieee} did not describe how or if they accounted for serial
dependence when applying the two-sample KS test.
The distribution of the test statistic under
the null hypothesis depends on the unknown parameters and unknown 
serial dependence. Therefore, the null distribution under dependence cannot be
used,
and the challenge is to approximate said distribution.
\citet{zeimbekakis2022misuses} proposed a semiparametric bootstrap
where the serial dependence structure is modeled via a working serial copula.

A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semiparametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for,
but this could affect results if the estimated serial
dependence is too far from the truth. For this reason, a nonparametric method
is motivated.

@@

First, let us review how the KS statistic is computed for an independent
sample with fitted parameters. Let $\hat\theta_n$ denote the parametrically
fitted parameters, which can be obtained from any consistent estimator with
asymptotic normality properties,  Let $F_n$ be the empirical distribution
function based on $X_1,...,X_n$.
Define 
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}
Then, the one-sample KS goodness of fit statistic is 
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta)|.
\end{equation*}

@@

In this simulation study, our objective is twofold: firstly, to demonstrate that
under the null distribution, our test rejects the null hypothesis ($H_0$) 
approximately at the specified significance level. Secondly, we aim to
illustrate that under the alternate distribution, our test reliably rejects
$H_0$ at a high frequency. The fulfillment of both these criteria would indicate
the method's efficacy.

\subsection{Size}
We first must evaluate whether this method works when $H_0$ is true. To
test this, we can
generate a simulated sample $X_t$ with a certain marginal distribution 
$F_\theta$,
and use our method to test if $X_t$ indeed has the marginal distribution $F$ 
with some unknown $\theta$. If the test holds its size, the 
p-value
of the test should be uniformly distributed between 0 and 1. We must try the
method with different marginal distributions to ensure that it is robust.
In order for the method to work, a large sample size may be necessary.

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
\jy{The language issues are more than I expected. Please do a careful proofread,
  now that you are more experienced.}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution and $\Phi$ is the distribution function of the standard normal
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are
\[
  \{-0.884, -0.686, -0.389, -0.033, 0.327, 0.639, 0.861\}.
\]

@@

Using R packages \textsl{qqplotr} and \textsl{ggplot2} \citep{qqplotr, ggplot2},
we constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, the
points will be aligned with the diagonal line. This 
indicates that under the null hypothesis, our method will only reject the $H_0$
at the rate of the significance level.
\jy{Describe the table not zoomed-in figure.}\mc{addressed}
A table of the rejection rates for each scheme
also provided.

\begin{figure}[tbp]
  \centering
  \includegraphics[width = \textwidth]{figures/normal}
  \caption{A Q-Q plot of the p-values testing that a distribution
  generated from a $N(8,8)$ data generating process is normal.}
  \label{fig:normal}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/gamma}
  \caption{A Q-Q plot of the p-values testing that a distribution
  generated from a $\gamma(8,1)$ data generating process is gamma distributed.}
  \label{fig:gamma}
\end{figure}

@@

From Figures~\ref{fig:normal} and \ref{fig:gamma}, we can observe that 
the extent to which the p-values are uniformly distributed
is dependent on the sample size and the level of serial 
dependence.
A small sample size like $n = 100$ or $200$ seems adequate when Kendall's
$\tau$ is lower than 0.25. For Kendall's $\tau \geq 0.5$, a sample larger than
$n = 200$ seems necessary for the p-values to be normally distributed. For
Kendall's $\tau \geq 0.75$, a sample larger than $n = 800$ appears to be 
necessary, as the p-values are not aligned with the line in any of the Q-Q 
plots. This
is not necessarily a cause for concern, as (for normal margins)
a Kendall's $\tau$ corresponds to
a $\phi$ of 0.924, which is very high.
Additionally, positive serial dependence appears to misalign the distribution
of the p-values more so than negative serial dependences of the same strength.
Results are very similar for normal and gamma margins.
\jy{Summarize findings in bullet points before forming text.}

From Tables~\ref{table:rr_norm} and \ref{table:rr_gamma}, we can observe 
that the agreement between empirical size (rejection rate)
and nominal size (the
significance level) improves as sample size increases, but for $\tau = -0.75$
and $\tau = 0.75$, a sample size larger than 800 seems to be necessary for that 
agreemet to be reached. Using the criteria of approximate agreement between 
empirical size and nominal size, we can say that our approach is working 
for $-0.25 < \tau < 0.25$ for $n > 100$. However a larger sample size seems 
necessary, greater than 400 perhaps, for $\tau$ as strong as $0.5$. To 
summarize, for weaker dependence, our method works without need for a 
particularly
large sample size, whereas for moderate dependence, a larger sample size may 
be required. Because $\tau$ is typically less than 0.5, it is still reasonable
to
recommend our method with a smaller sample size.

\subsection{Power}
The power of the proposed test is investigated with data generated from
distributions that are not in the hypothesized distribution family. If the test
is indeed powerful, we would ideally want the probability of
rejecting $H_0$ when $H_0$ is false\jy{beta has been
  used. Discuss in terms of power: the probability of rejecting $H_0$ when $H_0$
  is false.}\mc{addressed}
to be 1, but we
expect it to be generally high. Again, we must try the method with different
marginal distributions to ensure that it is robust.

\jy{Description too wordy, Be concise. Basically generate data from one
  distribution but test for the other distribution.}\mc{addressed}
We generated the same scheme of time series as we did when testing size, but
tested that the time series was from the other distribution.
Because the support of $N(8, 8)$ is
$(-\infty, \infty)$, but the support of $\Gamma(8, 1)$ is $(0, \infty)$, we used
the \textsl{truncdist} package \citep{truncdist} to truncate the series at 
values less than or equal to 0 when testing that $N(8, 8)$ sample was
from $\Gamma(8, 1)$. For each setting for $F$, $\tau$, and $n$, we replicate the 
method $1000$ times 
to obtain a p-value $p_r$ for each $r \in \{1, \ldots, 1000\}$.

\jy{alpha has been used as gamma parameters. This is not necessary, Just say we
  checked the power when the significance level is .05.}\mc{addressed}
We can compute a rejection rate 
at a nominal significance level of 0.05.
Additionally, we can construct a 95\% confidence interval for 
this rate. Below are tables containing rejection rates showcasing the 
differences when simulation settings are changed.

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/rr}
  \caption{A plot of the rejection rates as a function of $\tau$ for
 $n \in \{100, 200, 400, 800\}$ and marginal distribution 
 $\in \{N(8,8), \Gamma(8,1)\}$.}
  \label{fig:rr}
\end{figure}

Figure~\ref{fig:rr} displays the empirical power curve of the test at
significance level 0.05 as the lag-1 Kendall's~$\tau$ increases from
$-0.5$ to $0.5$. We omitted -0.75 and 0.75 from this study because we already 
found that our method doesn't work under the null hypothesis for such high
dependence levels. It appears that rejection rate seems to decline as $\tau$ 
increases. Using $n \geq 400$ seems to be good enough
if $\tau < 0.5$, but using a $n \geq 800$ will result in 
rejection rates close to 1 for $0.5 < \tau < 0.5$.
Results for $n \geq 400$
do not appear to be to different when comparing gamma and normal 
margins, indicating that the method is robust to the marginal distribution.
The rejection rates for $n = 800$ are very 
high (almost 1), indicating that under $H_A$, the test is very powerful 
(low probability of failed rejection $\beta$) when a large enough
sample size is used.

@@

\jy{Put them into one table; add results for paremetric bootstrap with
  Zeimbakakis' approach with serial dependence accounted for; label the methods
  as serial dependence accounted for and not accounted for, two methods
  (parametric and nonparametric) under each.}

\jy{Discuss the difference between the semiparametric bootstrap and the proposed
  nonparametric bootstrap. The semiparametric method depends on a working copula
  model for the serial dependence. If that is severely misspecified, I'd expect
  poor performance. The proposed method does not need to specify that dependence
  structure.}
\mc{addressed earlier}

@@

\jy{Stock returns are known to be serially dependent and heavy-tailed. Search
  google scholar ``stylized facts of daily returns series'' and cite them.}
\mc{addressed}

@@

We have also 
demonstrated possible applications of this method to financial data.
Through comparison with the nonparametric
bootstrap bias correction developed by \citet{babu2004goodness}
and the parametric and semiparametric
bootstrap methods developed by \citet{zeimbekakis2022misuses}, we have 
underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
\jy{Too vague.}\mc{addressed}
the method's performance on different marginal distributions such
as the generalized extreme value distribution and
different dependence structures such as MA, ARMA, and ARIMA processes. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.

\jy{summarize remarks in bullet points first. Let's talk on Tuesday.}

>>>>

\documentclass[12pt, titlepage, letterpaper]{article}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{setspace}
\allowdisplaybreaks

@@

\title{Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test}
  \author{Mathew Chandy \\
  % \href{mailto:mathew.chandy@uconn.edu}
  % {\nolinkurl{mathew.chandy@uconn.edu}}\\
 Elizabeth Schifano\\
 Jun Yan\\
 Xianyang Zhang
}

@@

Another prevalent misapplication of the KS test arises when data exhibit serial
dependence, which is often overlooked in analyses. For example, in an
investigation of performance variation in high-performance computing systems,
\citet{tuncer2019ieee} did not describe how or if they accounted for serial
dependence when applying the two-sample KS test.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent.
The distribution of the test statistic under the null hypothesis depends on both
the unknown parameters and unknown serial dependence. 
\citet{zeimbekakis2022misuses} recommend a semiparametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.
Nonetheless, if the working dependence model is too far from the truth, the
results of the tests would not be reliable. For this reason, a nonparametric
method that does not specify the serial dependence is desired.

@@

First, let us review how the KS statistic is computed for an independent
sample with fitted parameters. Let $\hat\theta_n$ denote the parametrically
fitted parameters, which can be obtained from any consistent estimator with
asymptotic normality properties,  Let $F_n$ be the empirical distribution
function based on $X_1,...,X_n$.
Define 
\begin{equation*}
Y_n(x; \hat\theta_n) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}
Then, the one-sample KS goodness of fit statistic is 
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta_n)|.
\end{equation*}

@@

In this simulation study, our objective is twofold: firstly, to demonstrate that
under the null distribution, our test rejects the null hypothesis ($H_0$) 
approximately at the specified size or significance level. Secondly, we aim to
illustrate that under the alternate distribution, our test rejects
$H_0$ with substantial power. The fulfillment of both criteria would indicate
the method's efficacy.

\subsection{Size}
To assess the actual size of the test, our strategy is to
generate a large number of stationary time series $X_t$ with a certain marginal
distribution $F(\cdot; \theta)$, and for each time series, use our method to
test if $X_t$ follows a hypothesized marginal distribution $F(\cdot; \theta)$
with some unknown $\theta$. If the test holds its size, the p-value
of the test should be uniformly distributed between 0 and 1. We will validate
method with different marginal distributions to ensure that it is robust.
In order for the method to work, a large number of replicates is necessary.

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with seven levels of Kendall's
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it is invariant to monotone marginal
transformations. To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive or AR(1) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
\jy{What was the value of $\sigma_\epsilon$?}
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution and $\Phi$ is the distribution function of the standard normal
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the lag-1 autocorrelations are
\[
  \{-0.884, -0.686, -0.389, -0.033, 0.327, 0.639, 0.861\}.
\]

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width = \textwidth]{figures/normal}
  \includegraphics[width = \textwidth]{figures/gamma}
  \caption{Q-Q plots of the p-values testing that a time series
    have marginal normal distribution with true data generating distribution
    being $N(8,8)$ (upper) and marginal gamma distribution with true data
    generating distribution $\Gamma(8,1)$ (lower)}
  \label{fig:normal}
\end{figure}

Figure~\ref{fig:qq} shows the Quantile-Quantile (Q-Q)
plots of the p-values testing the marginal distribution of the stationary time
series generated from the null hypothesis, created with R packages
\textsl{qqplotr} and \textsl{ggplot2} \citep{qqplotr, ggplot2}. If the
distribution of the p-values is uniform, the points will be aligned with the
diagonal line. The plots suggest that the extent to which the p-values are
uniformly distributed is dependent on the sample size and the level of serial 
dependence.
A small sample size like $n = 100$ or $200$ seems adequate when Kendall's
$\tau$ is lower than 0.25. For Kendall's $\tau \geq 0.5$, a sample larger than
$n = 200$ seems necessary for the p-values to be normally distributed. For
Kendall's $\tau \geq 0.75$, a sample larger than $n = 800$ appears to be 
necessary, as the p-values are not aligned with the line in any of the Q-Q 
plots. This
is not necessarily a cause for concern, as (for normal margins)
a Kendall's $\tau$ corresponds to
a $\phi$ of 0.924, which is very high.
Additionally, positive serial dependence appears to misalign the distribution
of the p-values more so than negative serial dependences of the same strength.
Results are very similar for normal and gamma margins.

@@

\jy{merge the two tables; use bottomrule and toprule instead of hline;
  put them in this file instead of inputing.}

For a closer scrutiny on whether the test holds its size at commonly used
significance levels 0.01, 0.05, and 0.10, Table~\ref{table:size}
summarizes the empirical sizes of the tests based on 1000 replicates.
The agreement between the empirical sizes and the nominal size
improves as sample size increases, as expected. For $\tau = -0.75$
and $\tau = 0.75$, a sample size as large as 800 is still not sufficient for the
agreemet to be reached. For weaker dependence levels, our approach is working
for $-0.25 < \tau < 0.25$ for $n = 100$. However, a larger sample size seems 
necessary, greater than 400 perhaps, for $\tau$ as strong as $0.5$. To 
summarize, for weaker dependence, our method works without need for a 
particularly large sample size, whereas for moderate dependence, a larger sample
size is required. For practical situations with a $\tau$ of magnitude 0.5 or
lower, it is still reasonable to recommend our method with $n \ge 400$.

\subsection{Power}
The power of the proposed test is investigated with data generated from
distributions that are not in the hypothesized distribution family. 
We generated the same scheme of time series as we did in the size investigation,
but tested that the time series was from the other distribution.
Because the support of $N(8, 8)$ is
$(-\infty, \infty)$, but the support of $\Gamma(8, 1)$ is $(0, \infty)$, we used
the \textsl{truncdist} package \citep{truncdist} to truncate the series at 
values less than or equal to 0 when testing that $N(8, 8)$ sample was
from $\Gamma(8, 1)$. For each setting for $F$, $\tau$, and $n$, we replicate the 
method $1000$ times 
to obtain a p-value $p_r$ for each $r \in \{1, \ldots, 1000\}$.
The empirical rejection rates were obtained at a nominal significance level 0.05.
Additionally, we constructed 95\% confidence intervals the actual rejection
rejection rate.

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/rr}
  \caption{Power curve as a function of $\tau$ for
    $n \in \{100, 200, 400, 800\}$ and marginal distribution 
    $\in \{N(8,8), \Gamma(8,1)\}$.\jy{clarify the data generating distribution and the
      hypothesized distribution}
  }
  \label{fig:rr}
\end{figure}

Figure~\ref{fig:rr} displays the empirical power curve of the test at
significance level 0.05 as the lag-1 Kendall's~$\tau$ increases from
$-0.5$ to $0.5$. We omitted -0.75 and 0.75 from this study because we already 
found that our method doesn't work under the null hypothesis for such high
dependence levels. It appears that rejection rate seems to decline as $\tau$
\jy{comment on the asymmetry of the power in response to tau.}
increases. Using $n \geq 400$ seems to be good enough
if $\tau < 0.5$, but using a $n \geq 800$ will result in 
rejection rates close to 1 for $0.5 < \tau < 0.5$.
Results for $n \geq 400$
do not appear to be to different when comparing gamma and normal 
margins, indicating that the method is robust to the marginal distribution.
The rejection rates for $n = 800$ are very 
high (almost 1), indicating that under $H_A$, the test is very powerful 
when a large enough sample size is used.

@@

@@

@@

We have also 
demonstrated possible applications of this method to financial data.
Through comparison with the nonparametric
bootstrap bias correction developed by \citet{babu2004goodness}
and the parametric and semiparametric
bootstrap methods developed by \citet{zeimbekakis2022misuses}, we have 
underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
the method's performance on different marginal distributions such
as the generalized extreme value distribution and
different dependence structures such as MA, ARMA, and ARIMA
processes. Additionally, one could propose a similar method
for comparing two serially dependent methods. One could also apply the method
showcased in this paper to fields like earthquake prediction and astronomy.

\jy{QC bib entries: page numbers missing for some journal articles.}
